Comparing human and model-based forecasts of COVID-19 in Germany and Poland

### 2022-05-30 manuscript/manuscript.tex
use PLOS latex template

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Comparing human and model-based forecasts of COVID-19 in Germany and Poland},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{lineno}

@@

\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Comparing human and model-based forecasts of COVID-19 in Germany and Poland}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\setstretch{2}
\emph{Nikos I. Bosse, Sam Abbott, Johannes Bracher, Habakuk Hain, Billy J. Quilty, Mark Jit, Centre for the Mathematical Modelling of Infectious Diseases COVID-19 Working Group, Edwin van Leeuwen, Anne Cori, Sebastian Funk}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Forecasts based on epidemiological modelling have played an important role in shaping public policy throughout the COVID-19 pandemic. This modelling combines knowledge about infectious disease dynamics with the subjective opinion of the researcher who develops and refines the model and often also adjusts model outputs. Developing a forecast model is difficult, resource- and time-consuming. It is therefore worth asking what modelling is able to add beyond the subjective opinion of the researcher alone. To investigate this, we analysed different real-time forecasts of cases of and deaths from COVID-19 in Germany and Poland over a 1-4 week horizon submitted to the German and Polish Forecast Hub. We compared crowd forecasts elicited from researchers and volunteers, against a) forecasts from two semi-mechanistic models based on common epidemiological assumptions and b) the ensemble of all other models submitted to the Forecast Hub. We found crowd forecasts, despite being overconfident, to outperform all other methods across all forecast horizons when forecasting cases (weighted interval score relative to the Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational models performed comparably better when predicting deaths (rel. WIS 1.26), suggesting that epidemiological modelling and human judgement can complement each other in important ways.

\hypertarget{author-summary}{%
\section{Author summary}\label{author-summary}}

Mathematical models of COVID-19 have played a key role in informing governments across the world. While mathematical models are informed by our knowledge of infectious disease dynamics, they are ultimately developed and iteratively adjusted by the researchers and shaped by their subjective opinions. To investigate what modelling is able to add beyond the subjective opinion of the researcher alone, we compared human forecasts with model-based predictions of COVID-19 cases and deaths submitted to the so-called German/Polish Forecast Hub (which collates a variety of models from a range of teams).\\
We found that our human forecasts consistently outperformed an aggregate of all available model-based forecasts when predicting cases, but not when predicting deaths. Our findings suggest that human insight may be most valuable when forecasting highly uncertain quantities, which depend on many factors that are hard to model using equations, while mathematical models may be most useful in settings like predicting deaths, where leading indicators with a clear connection to the target variable are available. This potentially has very relevant policy implications, as agencies informing policy-makers could benefit from routinely eliciting human forecasts in addition to model-based predictions to inform policies.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Infectious disease modelling has a long tradition and has helped inform public health decisions both through scenario modelling, as well as actual forecasts of (among others) influenza (e.g. 1,2--4), dengue fever (e.g. 5,6,7), ebola (e.g. 8,9), chikungunya (e.g. 10,11) and now COVID-19 (e.g. 12,13--17). Applications of epidemiological models differ in the way they make statements about the future. Forecasts aim to predict the future as it will occur, while scenario modelling and projections aim to represent what the future could look like under certain scenario assumptions or if conditions stayed the same as they were in the past. Forecasts can be judged by comparing them against observed data. Since it is much harder to fairly assess the accuracy and usefulness of projections and scenario modelling in the same way, this work focuses on forecasts, which represent only a subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been collected, aggregated and compared by Forecast Hubs such as the US Forecast Hub (13,14), the German and Polish Forecast Hub (15,16) and the European Forecast Hub (17). Often, different individual forecasts are combined into a single forecast, e.g.~by taking the mean or median of all forecasts. These ensemble forecasts usually tend to perform better and more consistently than individual forecasts (see e.g. (6); (18)).

@@

Previous work has examined such direct human forecasts in various contexts, such as geopolitics (19,20), meta-science (21,22), sports (23) and epidemiology (11,24,25). Several prediction platforms (26--28) and prediction markets (29) have been created to collate expert and non-expert predictions. However, with the notable exception of (11), these forecasts were not designed to be evaluated alongside model-based forecasts and usually follow their own (often binary) prediction formats. Direct human forecasts may be able to take into account insights and relationships between variables which are hard to specify using epidemiological models. However, it is not entirely clear in which situations human forecasts perform well or badly. For example, (11) found that humans could outperform computer models at predicting the 2014/15 and 2015/16 flu season in the US, a setting where the disease was well known and information about previous seasons was available. However, humans tended to do slightly worse at predicting the 2014/15 outbreak of chikungunya in the Americas, a disease previously largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts relative to model-based forecasts and discuss the added benefit of epidemiological modelling over human judgement alone. As a case study, we use different forecasts, involving varying degrees of human intervention, which we submitted in real time to the German and Polish Forecast Hub. In contrast to (11) we elicited not only point predictions, but full predictive distributions (``probabilistic forecasts'', see e.g. (30)) from participants. This allows us to compare not only predictive accuracy, but also how well human forecasters and model-based forecasts were able to quantify forecast uncertainty.

@@

We created and submitted the following forecasts to the German and Polish Forecast Hub: 1) a direct human forecast (henceforth called ``crowd forecast''), elicited from participants through a web application (31) and 2) two semi-mechanistic model-based forecasts (``renewal model'' and ``convolution model'') informed by basic assumptions about COVID-19 epidemiology. While the two semi-mechanistic forecasts were necessarily shaped by our implicit assumptions and decisions, they were designed such as to minimise the amount of human intervention involved. For example, we refrained from adjusting model outputs or refining the models based on past performance. Forecasts were created in real time over a period of 21 weeks from October 12th 2020 until March 1st 2021 and submitted to the German and Polish Forecast hub (15,16). All code and tools necessary to generate the forecasts and make a forecast submission are available in the \texttt{covid.german.forecasts} R package (32). This repository also contains a record of all forecasts submitted to the German and Polish Forecast Hub. Forecasts were evaluated using a variety of scoring metrics and compared among each other and against an ensemble of all other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the (17)) elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be made every Monday (with submissions allowed until Tuesday 3pm) and were permitted to use any data that was available by Monday 11.59pm. We submitted forecasts for incident and cumulative weekly reported numbers of cases of and deaths from COVID-19 on a national level in Germany and Poland over a one to four week forecast horizon. Forecasts were submitted on Mondays, but weeks were defined as ending on a Saturday (and starting on Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26 days. Submissions were required in a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub were combined into different ensembles every week, with the median ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the median of all submitted \(\alpha\)-quantiles) being the default ensemble shown on all official Forecast hub visualisations (\url{https://kitmetricslab.github.io/forecasthub/forecast}).

Data on daily reported test positive cases and deaths linked to COVID-19 were provided by the organisers of the German and Polish Forecast hub. Until December 14th, 2020, these data were sourced from the European Centre for Disease Control (33). After ECDC stopped publishing daily data, observations were sourced from the Robert Koch Institute (RKI) and the Polish Ministry of Health for the remainder of the submission period (34). These data are subject to reporting artefacts, (such as for example delayed case reporting in Poland on the 24th November, (35)), changes in reporting over time, and variation in testing regimes (for example in Germany from the 11th of November on, (36)). The ECDC data as well as the data published by the Polish Ministry of Health were also subject to data revisions, although most of them (with a notable exception of a data update for October 12 2020 in Germany) only affected daily, not weekly data (see Figures \ref{fig:daily-truth-update} and \ref{fig:weekly-truth-update}).

@@

Our crowd forecasts were created as an ensemble of forecasts made by individual participants every week through a web application (\url{https://cmmid-lshtm.shinyapps.io/crowd-forecast/}). Weekly forecasts had to be submitted before Tuesday 12pm every week, but participants were asked to only use any information or data that was already available by Monday night. The application was built using the \texttt{shiny} and \texttt{golem} R packages (37,38) and is available in the \texttt{crowdforecastr} R package (31). To make a forecast in the application participants could select a predictive distribution (with the default being log-normal) to represent the probability that the forecasted quantity took certain values. Median and width of the uncertainty could be adjusted by either interacting with a figure showing their forecast or providing numerical values (see screenshot in Figure \ref{fig:screenshot} in the SI). The default shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e.~as \(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)). A comparison of the crowd forecasts against the default baseline shown in the application is displayed in Figure \ref{fig:compare-forecasters} in the Appendix.
Our interface also allowed participants to view past observations based on the hub data, as well as their forecasts, on a logarithmic scale and presented additional contextual COVID-19 data sourced from (39). These data included, for example, notifications of both test positive COVID-19 cases and COVID-19 linked deaths and the number of COVID-19 tests conducted over time. From November 26 2020 on we displayed weekly small reports with a visualisation of past forecasts and scores on our website, epiforecasts.io.

@@

The study was approved by the Observational / Interventions Research Ethics Committee at the London School of Hygiene \& Tropical Medicine, LSHTM Ethics Reference: 22290.

@@

>>>>

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{lineno}
\usepackage[T1]{fontenc}
\linenumbers
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\usepackage{caption}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{London School of Hygiene \& Tropical Medicine,Karlsruhe
Institute of Technology Institute of Economic Theory and Statistics,Max
Planck Institute for Multidisciplinary Sciences,UK Health Security
Agency,Imperial College London}
}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Comparing human and model-based forecasts of COVID-19 in
Germany and
Poland} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nikos I. Bosse\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\textsuperscript{*},
Sam Abbott\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Johannes Bracher\textsuperscript{\getAff{Karlsruhe Institute of
Technology Institute of Economic Theory and Statistics}},
Habakuk Hain\textsuperscript{\getAff{Max Planck Institute for
Multidisciplinary Sciences}},
Billy J. Quilty\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Mark Jit\textsuperscript{\getAff{London School of Hygiene \& Tropical
Medicine}},
Centre for the Mathematical Modelling of Infectious Diseases COVID-19
Working Group\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}},
Edwin van Leeuwen\textsuperscript{\getAff{UK Health Security
Agency}, \getAff{London School of Hygiene \& Tropical Medicine}},
Anne Cori\textsuperscript{\getAff{Imperial College London}},
Sebastian Funk\textsuperscript{\getAff{London School of Hygiene \&
Tropical Medicine}}\\
\bigskip
\textbf{\getAff{London School of Hygiene \& Tropical Medicine}}London
School of Hygiene \& Tropical Medicine, Keppel Street, London WC1E7HT,
United Kingdom\\
\textbf{\getAff{Karlsruhe Institute of Technology Institute of Economic
Theory and Statistics}}Karlsruhe Institute of Technology - Institute of
Economic Theory and Statistics, Post Box 6980, 76049 Karlsruhe,
Germany\\
\textbf{\getAff{Max Planck Institute for Multidisciplinary Sciences}}Max
Planck Institute for Multidisciplinary Sciences, Am Faßberg 11, 37077
Göttingen, Germany\\
\textbf{\getAff{UK Health Security Agency}}UK Health Security Agency,
London NW9 5EQ, United Kingdom\\
\textbf{\getAff{Imperial College London}}Imperial College London,
Norfolk Place, London W2 1PG, United Kingdom\\
\bigskip
* Corresponding author: nikos.bosse@lshtm.ac.uk\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Forecasts based on epidemiological modelling have played an important
role in shaping public policy throughout the COVID-19 pandemic. This
modelling combines knowledge about infectious disease dynamics with the
subjective opinion of the researcher who develops and refines the model
and often also adjusts model outputs. Developing a forecast model is
difficult, resource- and time-consuming. It is therefore worth asking
what modelling is able to add beyond the subjective opinion of the
researcher alone. To investigate this, we analysed different real-time
forecasts of cases of and deaths from COVID-19 in Germany and Poland
over a 1-4 week horizon submitted to the German and Polish Forecast Hub.
We compared crowd forecasts elicited from researchers and volunteers,
against a) forecasts from two semi-mechanistic models based on common
epidemiological assumptions and b) the ensemble of all other models
submitted to the Forecast Hub. We found crowd forecasts, despite being
overconfident, to outperform all other methods across all forecast
horizons when forecasting cases (weighted interval score relative to the
Hub ensemble 2 weeks ahead: 0.89). Forecasts based on computational
models performed comparably better when predicting deaths (rel. WIS
1.26), suggesting that epidemiological modelling and human judgement can
complement each other in important ways.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.
\section*{Author summary}
Mathematical models of COVID-19 have played a key role in informing
governments across the world. While mathematical models are informed by
our knowledge of infectious disease dynamics, they are ultimately
developed and iteratively adjusted by the researchers and shaped by
their subjective opinions. To investigate what modelling is able to add
beyond the subjective opinion of the researcher alone, we compared human
forecasts with model-based predictions of COVID-19 cases and deaths
submitted to the so-called German/Polish Forecast Hub (which collates a
variety of models from a range of teams). \textbar{} We found that our
human forecasts consistently outperformed an aggregate of all available
model-based forecasts when predicting cases, but not when predicting
deaths. Our findings suggest that human insight may be most valuable
when forecasting highly uncertain quantities, which depend on many
factors that are hard to model using equations, while mathematical
models may be most useful in settings like predicting deaths, where
leading indicators with a clear connection to the target variable are
available. This potentially has very relevant policy implications, as
agencies informing policy-makers could benefit from routinely eliciting
human forecasts in addition to model-based predictions to inform
policies.

@@

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Infectious disease modelling has a long tradition and has helped inform
public health decisions both through scenario modelling, as well as
actual forecasts of (among others) influenza {[}e.g. 1,2--4{]}, dengue
fever {[}e.g. 5,6,7{]}, ebola {[}e.g. 8,9{]}, chikungunya {[}e.g.
10,11{]} and now COVID-19 {[}e.g. 12,13--17{]}. Applications of
epidemiological models differ in the way they make statements about the
future. Forecasts aim to predict the future as it will occur, while
scenario modelling and projections aim to represent what the future
could look like under certain scenario assumptions or if conditions
stayed the same as they were in the past. Forecasts can be judged by
comparing them against observed data. Since it is much harder to fairly
assess the accuracy and usefulness of projections and scenario modelling
in the same way, this work focuses on forecasts, which represent only a
subset of all epidemiological modelling.

Since March 2020, forecasts of COVID-19 from multiple teams have been
collected, aggregated and compared by Forecast Hubs such as the US
Forecast Hub {[}13,14{]}, the German and Polish Forecast Hub {[}15,16{]}
and the European Forecast Hub {[}17{]}. Often, different individual
forecasts are combined into a single forecast, e.g.~by taking the mean
or median of all forecasts. These ensemble forecasts usually tend to
perform better and more consistently than individual forecasts (see e.g.
{[}6{]}; {[}18{]}).

@@

Previous work has examined such direct human forecasts in various
contexts, such as geopolitics {[}19,20{]}, meta-science {[}21,22{]},
sports {[}23{]} and epidemiology {[}11,24,25{]}. Several prediction
platforms {[}26--28{]} and prediction markets {[}29{]} have been created
to collate expert and non-expert predictions. However, with the notable
exception of {[}11{]}, these forecasts were not designed to be evaluated
alongside model-based forecasts and usually follow their own (often
binary) prediction formats. Direct human forecasts may be able to take
into account insights and relationships between variables which are hard
to specify using epidemiological models. However, it is not entirely
clear in which situations human forecasts perform well or badly. For
example, {[}11{]} found that humans could outperform computer models at
predicting the 2014/15 and 2015/16 flu season in the US, a setting where
the disease was well known and information about previous seasons was
available. However, humans tended to do slightly worse at predicting the
2014/15 outbreak of chikungunya in the Americas, a disease previously
largely unobserved and unknown in these regions at the time.

In this study, we analyse the performance of direct human forecasts
relative to model-based forecasts and discuss the added benefit of
epidemiological modelling over human judgement alone. As a case study,
we use different forecasts, involving varying degrees of human
intervention, which we submitted in real time to the German and Polish
Forecast Hub. In contrast to {[}11{]} we elicited not only point
predictions, but full predictive distributions (``probabilistic
forecasts'', see e.g. {[}30{]}) from participants. This allows us to
compare not only predictive accuracy, but also how well human
forecasters and model-based forecasts were able to quantify forecast
uncertainty.

@@

We created and submitted the following forecasts to the German and
Polish Forecast Hub: 1) a direct human forecast (henceforth called
``crowd forecast''), elicited from participants through a web
application {[}31{]} and 2) two semi-mechanistic model-based forecasts
(``renewal model'' and ``convolution model'') informed by basic
assumptions about COVID-19 epidemiology. While the two semi-mechanistic
forecasts were necessarily shaped by our implicit assumptions and
decisions, they were designed such as to minimise the amount of human
intervention involved. For example, we refrained from adjusting model
outputs or refining the models based on past performance. Forecasts were
created in real time over a period of 21 weeks from October 12th 2020
until March 1st 2021 and submitted to the German and Polish Forecast hub
{[}15,16{]}. All code and tools necessary to generate the forecasts and
make a forecast submission are available in the
\texttt{covid.german.forecasts} R package {[}32{]}. This repository also
contains a record of all forecasts submitted to the German and Polish
Forecast Hub. Forecasts were evaluated using a variety of scoring
metrics and compared among each other and against an ensemble of all
other models submitted to the German and Polish Forecast Hub.

@@

The German and Polish Forecast Hub (now mostly merged into the {[}17{]})
elicits predictions for various COVID-19 related forecast targets from
different research groups every week. Forecasts had to be made every
Monday (with submissions allowed until Tuesday 3pm) and were permitted
to use any data that was available by Monday 11.59pm. We submitted
forecasts for incident and cumulative weekly reported numbers of cases
of and deaths from COVID-19 on a national level in Germany and Poland
over a one to four week forecast horizon. Forecasts were submitted on
Mondays, but weeks were defined as ending on a Saturday (and starting on
Sunday), meaning that forecast horizons were in fact 5, 12, 19 and 26
days. Submissions were required in a quantile-based format with 23
quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10,
0.15, \ldots, 0.95, 0.975, 0.99. Forecasts submitted to the Forecast Hub
were combined into different ensembles every week, with the median
ensemble (i.e., the \(\alpha\)-quantile of the ensemble is given by the
median of all submitted \(\alpha\)-quantiles) being the default ensemble
shown on all official Forecast hub visualisations
(https://kitmetricslab.github.io/forecasthub/forecast).

Data on daily reported test positive cases and deaths linked to COVID-19
were provided by the organisers of the German and Polish Forecast hub.
Until December 14th, 2020, these data were sourced from the European
Centre for Disease Control {[}33{]}. After ECDC stopped publishing daily
data, observations were sourced from the Robert Koch Institute (RKI) and
the Polish Ministry of Health for the remainder of the submission period
{[}34{]}. These data are subject to reporting artefacts, (such as for
example delayed case reporting in Poland on the 24th November,
{[}35{]}), changes in reporting over time, and variation in testing
regimes (for example in Germany from the 11th of November on, {[}36{]}).
The ECDC data as well as the data published by the Polish Ministry of
Health were also subject to data revisions, although most of them (with
a notable exception of a data update for October 12 2020 in Germany)
only affected daily, not weekly data (see Figures
\ref{fig:daily-truth-update} and \ref{fig:weekly-truth-update}).

@@

Our crowd forecasts were created as an ensemble of forecasts made by
individual participants every week through a web application
(https://cmmid-lshtm.shinyapps.io/crowd-forecast/). Weekly forecasts had
to be submitted before Tuesday 12pm every week, but participants were
asked to only use any information or data that was already available by
Monday night. The application was built using the \texttt{shiny} and
\texttt{golem} R packages {[}37,38{]} and is available in the
\texttt{crowdforecastr} R package {[}31{]}. To make a forecast in the
application participants could select a predictive distribution (with
the default being log-normal) to represent the probability that the
forecasted quantity took certain values. Median and width of the
uncertainty could be adjusted by either interacting with a figure
showing their forecast or providing numerical values (see screenshot in
Figure \ref{fig:screenshot} in the SI). The default shown was a
repetition of the last known observation with constant uncertainty
around it computed as the standard deviation of the last four changes in
weekly log observed forecasts (i.e.~as
\(\sigma(log(value4) - log(value3), log(value3) - log(value2), \ldots )\)).
A comparison of the crowd forecasts against the default baseline shown
in the application is displayed in Figure \ref{fig:compare-forecasters}
in the Appendix. Our interface also allowed participants to view past
observations based on the hub data, as well as their forecasts, on a
logarithmic scale and presented additional contextual COVID-19 data
sourced from {[}39{]}. These data included, for example, notifications
of both test positive COVID-19 cases and COVID-19 linked deaths and the
number of COVID-19 tests conducted over time. From November 26 2020 on
we displayed weekly small reports with a visualisation of past forecasts
and scores on our website, epiforecasts.io.

@@

The study was approved by the Observational / Interventions Research
Ethics Committee at the London School of Hygiene \& Tropical Medicine,
LSHTM Ethics Reference: 22290. Formal consent from participants was
taken in written form through the web application.

@@

\nolinenumbers
